Supplement: Bootstrapping Domain Specific Languages for Bayesian Program Learning

### 2018-02-08 paper.tex
text baseline data
supplement improvements

We compare  with four baselines on  held-out tasks:
\\\noindent \textbf{Ours--NN}, which lesions the recognition model.
\\\noindent \textbf{RF/DC}, which holds the generative model $(\mathcal{D},\theta)$ fixed and learns
a recognition model only from samples from the fixed generative model.
This is equivalent to our algorithm with $\lambda = \infty$ (see Section~\ref{grammarInductionSection})
and $\mathcal{L}_{\text{RM}} = \mathcal{L}_{\text{HM}}$ (see Section~\ref{recognitionSection}).
We call this baseline RF/DC because
this setup is closest to how RobustFill~\cite{devlin2017robustfill} and DeepCoder~\cite{balog2016deepcoder} are trained.
\\\noindent \textbf{PCFG}, which lesions the recognition model, fixes $\mathcal{D}$, and learns only $\theta$.
This is like learning the parameters of a PCFG.
\\\noindent \textbf{Enum}, which does no learning and just enumerates a frontier.
This is equivalent to the first wake cycle of our \system.

@@

  \\\midrule\multicolumn{6}{c}{\emph{String Editing}}\\\midrule
  \% solved&99\% &82\% \\
  Solve time&  0.9s&2.2s
  \\\bottomrule
  \end{tabular}
\caption{\% solved: fraction solved with 5 sec timeout. Solve time: averaged over solved tasks. MDL: $-\expect\left[\probability[x|p] \right]$. For domains other than symbolic regression MDL is 0 nats. For symbolic regression MDL is a proxy for \# of continuous degrees of freedom.}\label{baselineComparisons}  \end{table}

>>>>

We compare  with four baselines on  held-out tasks:
\\\noindent \textbf{Ours--NN}, which lesions the recognition model.
\\\noindent \textbf{RF/DC}, which holds the generative model $(\mathcal{D},\theta)$ fixed and learns
a recognition model only from samples from the fixed generative model.
This is equivalent to our algorithm with $\lambda = \infty$ (see Section~\ref{grammarInductionSection})
and $\mathcal{L}_{\text{RM}} = \mathcal{L}_{\text{HM}}$ (see Section~\ref{recognitionSection}).
We call this baseline RF/DC because
this setup is closest to how RobustFill~\cite{devlin2017robustfill} and DeepCoder~\cite{balog2016deepcoder} are trained.
\\\noindent \textbf{PCFG}, which lesions the recognition model, learns $\theta$, and fixes $\mathcal{D}$.
This is equivalent to our algorithm with $q(x) = \theta$ and $\lambda = \infty$,
and is like learning the parameters of a PCFG while not learning any of the structure.
\\\noindent \textbf{Enum}, which does no learning and just enumerates a frontier.
This is equivalent to the first wake cycle of \system.

@@

  \\\midrule\multicolumn{6}{c}{\emph{String Editing}}\\\midrule
  \% solved&99\% &82\% &9\%&24\%&18\%\\
  Solve time&  0.9s&2.2s&2.7s&2.3s&2.4s
  \\\bottomrule
  \end{tabular}
\caption{\% solved: fraction solved with 5 sec timeout. Solve time: averaged over solved tasks. MDL: $-\expect\left[\probability[x|p] \right]$. For domains other than symbolic regression MDL is 0 nats. For symbolic regression MDL is a proxy for \# of continuous degrees of freedom.}\label{baselineComparisons}  \end{table}

### 2018-02-08 supplement.tex
supplement improvements

\newcommand{\system}{\textsc{HelmholtzHacker}~}

@@

% The \icmltitle you define below is probably too long as a header.
% Therefore, a short form for the running title is supplied here:
\icmltitlerunning{Inducing Domain Specific Languages for Bayesian Program Learning}

@@

\twocolumn[
\icmltitle{Supplement to: Inducing Domain Specific Languages for Bayesian Program Learning}

@@

@@

Differentiate with respect to $\theta_e$ and set to zero 
\begin{align}
  \frac{c(x)}{\theta^{(x)}} &= N\frac{a(x)}{\sum_y a(y)\theta_y}
\end{align}
This equality holds if $\theta^{(x)} = c(x)/a(x)$:
\begin{align}
  \frac{c(x)}{\theta_x} &= a(x).\\
N\frac{a(x)}{\sum_y a(y)\theta_y}& = N\frac{a(x)}{\sum_y c(y)}   = N\frac{a(x)}{N} = a(x).
\end{align}
If this equality holds then $\theta_x \propto c(x)/a(x)$:
\begin{align}
  \theta_x = \frac{c(x)}{a(x)}\times \underbrace{\frac{\sum_y a(y)\theta_y}{N}}_{\text{Independent of $x$}}.
\end{align}

@@

>>>>

\newcommand{\system}{\textsc{DreamCoder}~}

@@

% The \icmltitle you define below is probably too long as a header.
% Therefore, a short form for the running title is supplied here:
\icmltitlerunning{Bootstrapping Domain Specific Languages for Bayesian Program Learning}

@@

\twocolumn[
\icmltitle{Supplement: Bootstrapping Domain Specific Languages for Bayesian Program Learning}

@@

If we were to maximize the ELBO bound to perform inference in
our generative model,
then, during DSL induction, we would seek a new $(\mathcal{D}^*,\theta^*)$ maximizing the following lower bound on the likelihood (along an unimportant regularizing term on the DSL):
\begin{align}
  &  \sum_{x\in X}\expect_{p\sim Q_x}\left[\log \probability[p|\mathcal{D}^*,\theta^*] \right] \label{EM}\\
  & Q_x(p)\triangleq \probability[p|x,\mathcal{D},\theta]
\end{align}
where $(\mathcal{D},\theta)$ is our current estimate of the generative
model.  These equations fall out of an EM-style derivation, and one
could replace $Q_x(p)$ with the recognition model $q(p|x)$, either
using importance sampling (so the expectation in~\ref{EM} is taken
over $q$ and we reweight using $Q_x$) or by directly using $q$ as our approximate posterior over the program that solves task $x$.

How the recognition model is worked into this is unimportant to the reason why we do not maximize a bound of this form:
this bound takes an expectation over the \emph{previous} sleep cycle's
posterior over programs,
and so the approximate posterior $Q_x$ at the next iteration
ends up being very close to previous approximate posterior.
Intuitively,
we want the DSL induction to be a function \emph{only} of the programs that we have found,
and \emph{not} be a function of how the previous generative model weighted them.
In practice,
we found that maximizing EM-style bounds, like the ELBO,
leads to a kind of hysteresis effect,
where the next generative model too closely matches the previous one,
causing the algorithm to easily become trapped in local optima.

@@

Differentiate with respect to $\theta_e$ and set to zero 
\begin{align}
  \frac{c(x)}{\theta_x} &= N\frac{a(x)}{\sum_y a(y)\theta_y}
\end{align}
This equality holds if $\theta_x = c(x)/a(x)$:
\begin{align}
  \frac{c(x)}{\theta_x} &= a(x).\\
N\frac{a(x)}{\sum_y a(y)\theta_y}& = N\frac{a(x)}{\sum_y c(y)}   = N\frac{a(x)}{N} = a(x).
\end{align}
If this equality holds then $\theta_x \propto c(x)/a(x)$:
\begin{align}
  \theta_x = \frac{c(x)}{a(x)}\times \underbrace{\frac{\sum_y a(y)\theta_y}{N}}_{\text{Independent of $x$}}.
\end{align}

@@

When is this lower bound tight? This lower bound is tight whenever all
of the types of the expressions in the DSL are not polymorphic, in which case our DSL is equivalent to a PCFG
and this estimator is equivalent to the inside/outside algorithm.
Polymorphism introduces context-sensitivity to the DSL,
and exactly maximizing the likelihood with respect to $\theta$
becomes intractable,
so for domains with polymorphic types we use this estimator.
